Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down

### 2023-04-04 paper/ms.tex
Small changes from ref report

\section{Introduction}
\label{sec:intro}
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fig/mockHST.pdf}
    \caption{Synthetic Hubble Space Telescope images of our simulations with and
    without the interstellar medium. The upper panels show an \Nbody{} only
    simulation while the lower panels show a simulation which includes the
    \SMUGGLE{} model for the interstellar medium. Each panel is $20\,\textrm{kpc}$
    to a side. Columns show different points in time, separated by
    $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
    length and strength. In the \SMUGGLE{} run, the bar remains at approximately
    the same length and strength over the course of the simulation. The \Nbody{}
    model is identical to the GALAKOS model, discussed in the text. Details on
    the production of these synthetic images is given in Appendix~\ref{app:hst}.}\label{fig:overview}
\end{figure*}

Approximately two-thirds of spiral disks host an elongated bar-like feature at
their centers \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our
own Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of
bars is not difficult to explain, since stellar disks simulated in isolation
almost always form bar-like structures \citep{1971ApJ...168..343H}. Several
studies have shown that a spherical potential, such as a stellar bulge or dark
matter halo, acts to stabilize stellar disks against bar formation
\citep[e.g.,][]{1973ApJ...186..467O, 1976AJ.....81...30H}.

@@

While the interaction between the bar and a dark matter halo is well-understood
theoretically, this is not the case for the interaction between a bar and a
gaseous disk. Some argue that the gas disk should slow down the bar more
\citep{2003MNRAS.341.1179A}, while others argue that the tendency of the bar to
drive gas inwards means the bar should speed up due to the effect of the gas
disk \citep{2013MNRAS.429.1949A, 2014MNRAS.438L..81A}. Since the gas phase
typically contributes only about $10-20\%$ of the mass of a galaxy at the
present day, one might naively expect it to have a subdominant effect on the
bar. However, because gas is collisional, it can participate in non-resonant
angular momentum exchange with the bar \citep{2011MNRAS.415.1027H}. Thus,
numerical work has shown that the gas phase can have a stronger influence on a
bar than its contribution to the mass of a galaxy would suggest
\citep{2010ApJ...719.1470V, 2013MNRAS.429.1949A}.

@@

We have performed a simulation of a disk galaxy using the finite-volume,
gravito-hydrodynamics code \AREPO{} \citep{2010MNRAS.401..791S}. We use the
galaxy formation model Stars and MUltiphase Gas in GaLaxiEs
\citep[\SMUGGLE{};][]{2019MNRAS.489.4233M}. This disk galaxy exhibits almost no
evolution in the bar pattern speed over several Gyr when the gas phase is
accounted for and properly modeled. This behavior has been observed in a few
previous works \citep{1993AA...268...65F, 2007ApJ...666..189B,
2009ApJ...707..218V, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}.

@@

The implications of our findings are numerous. First, we naturally explain why
nearly all observed galaxies are fast rotators without requiring the inner
regions of dark matter halos to be underdense \citep{1998ApJ...493L...5D,
2000ApJ...543..704D} or requiring new physics \citep{2021MNRAS.503.2833R,
2021MNRAS.508..926R}. Second, we show that the role of gas is of paramount
importance in studies which attempt to uncover the nature of dark matter from
its effect of slowing down the bar \citep{2021MNRAS.500.4710C,
2021MNRAS.505.2412C}. Third, we provide an explanation for how the Milky Way's
bar could be both long-lived and a fast rotator, of which there is some
observational evidence \citep{2019MNRAS.490.4740B}. And finally, we complicate
the picture of radial mixing expected to sculpt the Milky Way's disk
\citep{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
the pattern speed of the bar to change with time Our work does not alter
expectations for radial mixing induced by spiral arms
\citep{2002MNRAS.336..785S}.

@@

We performed an experiment to test this hypothesis by forcing the stellar disk
in the \SMUGGLE{} run to rotate at a constant angular rate and measuring the
torque on the bar by the gas phase at different rotation rates. The result of
this experiment is illustrated in Fig.~\ref{fig:equil}, which shows that a more
slowly rotating bar experiences a larger positive torque from the gas.

>>>>

\section{Introduction}
\label{sec:intro}
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fig/mockHST.pdf}
    \caption{Synthetic Hubble Space Telescope images of our simulations with and
    without the interstellar medium. The upper panels show an \Nbody{} only
    simulation while the lower panels show a simulation which includes the
    \SMUGGLE{} model for the interstellar medium. Each panel is
    $20\,\textrm{kpc}$ to a side. Columns show different points in time,
    separated by $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar
    grows in length and strength. In the \SMUGGLE{} run, the bar remains at
    approximately the same length and strength over the course of the
    simulation. The \Nbody{} model is identical to the GALAKOS model {\bf (with
    different particle numbers and softening lengths)}, discussed in the text.
    Details on the production of these synthetic images is given in
    Appendix~\ref{app:hst}.}\label{fig:overview}
\end{figure*}

Approximately two-thirds of spiral disks host an elongated bar-like feature at
their centers \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our
own Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of
bars is not difficult to explain, since stellar disks simulated in isolation
almost always form bar-like structures \citep{1971ApJ...168..343H}. Several
studies have shown that a {\bf hot, centrally concentrated mass distribution},
such as a stellar bulge or dark matter halo, acts to stabilize stellar disks
against bar formation \citep[e.g.,][]{1973ApJ...186..467O, 1976AJ.....81...30H}.

@@

{\bf While the fact that a bar is slowed down by a dark matter halo is
well-understood theoretically,} this is not the case for the interaction between
a bar and a gaseous disk. Some argue that the gas disk should slow down the bar
more \citep{2003MNRAS.341.1179A}, while others argue that the tendency of the
bar to drive gas inwards means the bar should speed up due to the effect of the
gas disk \citep{2013MNRAS.429.1949A, 2014MNRAS.438L..81A}. Since the gas phase
typically contributes only about $10-20\%$ of the mass of a galaxy at the
present day, one might naively expect it to have a subdominant effect on the
bar. However, because gas is collisional, it can participate in non-resonant
angular momentum exchange with the bar \citep{2011MNRAS.415.1027H}. Thus,
numerical work has shown that the gas phase can have a stronger influence on a
bar than its contribution to the mass of a galaxy would suggest
\citep{2010ApJ...719.1470V, 2013MNRAS.429.1949A}.

@@

We have performed a simulation of a disk galaxy using the finite-volume,
gravito-hydrodynamics code \AREPO{} \citep{2010MNRAS.401..791S}. We use the
galaxy formation model Stars and MUltiphase Gas in GaLaxiEs
\citep[\SMUGGLE{};][]{2019MNRAS.489.4233M}. This disk galaxy exhibits almost no
evolution in the bar pattern speed over several Gyr when the gas phase is
accounted for and {\bf robustly} modeled. This behavior has been observed in a
few previous works \citep{1993AA...268...65F, 2007ApJ...666..189B,
2009ApJ...707..218V, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}.

@@

The implications of our findings are numerous. First, we naturally explain why
nearly all observed galaxies are fast rotators without requiring the inner
regions of dark matter halos to be underdense \citep{1998ApJ...493L...5D,
2000ApJ...543..704D} or requiring new physics \citep{2021MNRAS.503.2833R,
2021MNRAS.508..926R}. Second, we show that the role of gas is of paramount
importance in studies which attempt to uncover the nature of dark matter from
its effect of slowing down the bar \citep{2021MNRAS.500.4710C,
2021MNRAS.505.2412C}. Third, we provide an explanation for how the Milky Way's
bar could be both long-lived and a fast rotator, of which there is some
observational evidence \citep{2019MNRAS.490.4740B}. And finally, we complicate
the picture of radial mixing expected to sculpt the Milky Way's disk
\citep{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
the pattern speed of the bar to change with time. Our work does not alter
expectations for radial mixing induced by spiral arms
\citep{2002MNRAS.336..785S}.

@@

{\bf We performed an experiment to test this hypothesis by freezing the stellar
disk in the \SMUGGLE{} run and forcing it to rotate at a constant angular rate.
This has the effect of forcing the bar to rotate as a solid body at a constant
angular rate which we control. The gas is evolved self-consistently with this
rotating disk.} We measured the torque on the bar by the gas phase at different
rotation rates. The result of this experiment is illustrated in
Fig.~\ref{fig:equil}, which shows that a more slowly rotating bar experiences a
larger positive torque from the gas.
